---
title: "Distributions of Catch by Species in MRIP Data"
author: "Ben Williams"
date: "May 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##R Markdown

```{r load packages}
library(tidyverse)
library(ggthemes)
```

The data, make sure psu_id is a character

```{r}
mrip <- read_csv("data/mrip_tidy2_17.csv",
                 col_types = cols(psu_id = col_character()))

```

What fish are most often caught?

Get a named vector of the sum of all the fish caught per species
```{r}
fish_nums <- mrip %>% 
  dplyr::select(ends_with("claim")) %>% 
  colSums() 
```
Make it a tibble
```{r}
fish_numbers <- tibble(spec = names(fish_nums),
                       number = unname(fish_nums))
```

plot the total catch by species
```{r}
fish_numbers %>% 
  ggplot(aes(spec,number))+
  geom_col()+
  coord_flip()
```

That is a bit too hard to decipher, let's keep only species caught at least 100 times in 2017
```{r}
fish_numbers %>% 
  filter(number >= 100) %>% 
  ggplot(aes(reorder(spec,number),number,fill=spec))+
  geom_col(show.legend = F)+
  coord_flip()
```



Let's look at the distribution of Red Snapper in the 2017 MRIP data

```{r}
mrip %>% 
  ggplot(aes(`RED SNAPPER_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP Red Snapper Catch")
```

What does the plot look like without *zero's*

```{r}
mrip %>% 
  filter(`RED SNAPPER_claim` > 0) %>% 
  ggplot(aes(`RED SNAPPER_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP Red Snapper Catch > 0")
```

Now Vermillion Snapper
```{r}
mrip %>% 
  ggplot(aes(`VERMILION SNAPPER_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP Vermillion Snapper Catch")
```

Now,exclude the *zero's*

```{r}
mrip %>%
  filter(`VERMILION SNAPPER_claim` > 0) %>% 
  ggplot(aes(`VERMILION SNAPPER_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP Vermillion Snapper Catch > 0")
```

Lane Snapper
```{r}
mrip %>% 
  ggplot(aes(`LANE SNAPPER_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP Lane Snapper Catch")
```

Now,exclude the *zero's*

```{r}
mrip %>%
  filter(`LANE SNAPPER_claim` > 0) %>% 
  ggplot(aes(`LANE SNAPPER_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP Vermillion Snapper Catch > 0")
```

White Grunt
```{r}
mrip %>% 
  ggplot(aes(`WHITE GRUNT_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP WHITE GRUNT Catch")
```

Now,exclude the *zero's*

```{r}
mrip %>%
  filter(`WHITE GRUNT_claim` > 0) %>% 
  ggplot(aes(`WHITE GRUNT_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP White Grunt Catch > 0")
```

Spanish Mackerel
```{r}
mrip %>% 
  ggplot(aes(`SPANISH MACKEREL_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP SPANISH MACKEREL Catch")
```

Now,exclude the *zero's*

```{r}
mrip %>%
  filter(`SPANISH MACKEREL_claim` > 0) %>% 
  ggplot(aes(`SPANISH MACKEREL_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP SPANISH MACKEREL Catch > 0")
```

White Grunt
```{r}
mrip %>% 
  ggplot(aes(`RED PORGY_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP RED PORGY Catch")
```

Now,exclude the *zero's*

```{r}
mrip %>%
  filter(`RED PORGY_claim` > 0) %>% 
  ggplot(aes(`RED PORGY_claim`)) +
  geom_histogram(bins=20) +
  ggtitle("2017 MRIP RED PORGY Catch > 0")
```

